Data cleaning for mass mobiliziation data version

### 2024-01-19 Scripts/authoritarian_only.Rmd
New draft post-AJPS reject

On August 9, 2020, Belarus held an election in which President Aleksander Lukashenko ran for a sixth term. The official results credited Lukashenko with over 80\% of the vote, a result which was regarded by observers as wildly inflated [@GolosPlatformPresents2020]. In response, hundreds of thousands of Belarusians took to the streets over the course of several months. Though the protests put significant strain on the regime, a harsh campaign of repression ultimately quelled the protests and Lukashenko was inaugurated into a new five-year term. The Belarusian case offers a useful illustration of a central puzzle in the study of election manipulation: why did the risk of costly protest fail to deter the incumbent from engaging in such widespread fraud? And why did the regime use fraud to claim such an implausibly high margin of victory?

@@

Testing this proposition requires looking beyond binary measures of protest, and toward measures of its scope. Research on the risk posed to non-democratic governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of broad-based protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010]. As Brancati [-@Brancati2016] finds, approximately two-thirds of pro-democracy electoral protests ended within three days, posing little threat to incumbents. This implicit assumption has largely been overlooked in theories and empirical studies of the relationship between election manipulation and protest, however. In most cases, the risk of protest has both been conceptualized and operationalized as a binary outcome: protests occur or they do not. This approach obscures the importance of the number and size of electoral protests in a ruling party's calculation of the risk involved in manipulation an election. Small, narrow protests may be easily contained by a well-resourced government, while more numerous, larger protest events pose a more significant challenge.

Using cross-national data on 1110 elections from 1989 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Vioence dataset [@Daxecker2019], this paper finds little support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs. Furthermore, it shows that more severe election fraud is not associated with more numerous or larger post-election protests. Instead, these protest dynamics appear driven more by structural and institutional factors, including the size of the political opposition and whether or not the incumbent chief executive is term limited.

Since fraud is not associated with the overall number or size of of events in a protest wave, the risk to governments has likely been overemphasized. That is, even if protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them. Once protests begin, the noisy signal of regime strength provided by the election is replaced by a very clear signal: the force of the regime's response to the emerging protest movement in contest with the mobilizational capacity of the opposition.  To the extent that larger protest waves are more damaging to the government than small ones, protest risk does not appear to be a substantial constraint on the severity of election fraud. Troublingly, these findings indicate that ruling parties should follow an 'all-in' strategy on election manipulation. More manipulation is not associated with a larger number of protests, but a large margin of victory is associated with a lower risk of protest initiation. These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory. They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations.

# Literature review

Broadly speaking, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework). The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs *may* manifest as protest, but may also manifest as a reduction in general cooperation with the regime. However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can *induce* cooperative behavior from some agents], factor which affects the demand-side cost of manipulation in most models. This insight can be found, in related forms, in both the literature on self-enforcing democracy and on the conduct of non-democratic elections. In the former literature, generally free and fair elections are thought to be a self-enforcing equilibrium through the threat of mass protest by wronged groups if the incumbent government rigs the results. In many formal models, including those discussed below, the decision to protest or consent to the election is the play available to citizens or opposition parties; protesting imposes a cost on the regime and/or carries a risk that the incumbent is overthrown.

In the major formulations of this argument, such as Fearon  [-@fearon_self-enforcing_2011] and Przeworski [-@przeworskiSelfenforcingDemocracy2006], 'democracy' is understood in procedural terms. For Przeworski, democracy requires that "authorization to rule is derived from elections" which designate winners and losers according to a set of rules [@przeworskiSelfenforcingDemocracy2006, p. 312]. Fearon adopts a similar definition for electoral democracy [@fearon_self-enforcing_2011, p. 1662]. Such arrangements are not limited to liberal democracies, and in fact constitute the large majority of regimes today. Many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020]\textemdash but this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.

There are thought to be two mechanisms by which election manipulation leads to protest, and thereby might restrain manipulation. The first is through revealing information about the strength of the government or ruling party. This information, in turn, feeds into strategic decision-making by opposition groups. This logic is often seen in formal models [@Egorov2021;@chernykh2015third]. Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @fearon_self-enforcing_2011;@Luo2020;@Rozenas2016b]. The risk of mass protest in these models is assumed to constitute a significant guardrail for election integrity in democracies, and a significant challenge for ruling parties in non-democracies. Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Little2012;@Hyde2014;@daxecker2012], civil society groups [@birch2017getting], and courts [@chernykh2015third]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance. Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons and across time\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.

Despite this consensus in the formal literature, there is relatively little empirical evidence to support the claim that manipulation per se indicates weakness, and thus increases protest risk. For example, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that post-election protest was not associated with the severity of fraud using several measures.^[It should be noted that Lankina and Skovoroda [-@Lankina2017] show that a different measure of fraud was associated with post-election protest in the subsequent presidential election.] In a cross-national study, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e. whether opponents are allowed to run, whether the media environment is pro-regime), rather than by the election-day manipulation that usually figures in theories of election integrity and protest. This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].

The literature on protest as a deterrent to manipulation also often overlooks the functions that elections, even highly controlled elections, may serve in non-democratic settings. For instance, Simpser [-@simpser2013] shows that high levels of manipulation can actually induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength. Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with costlier forms of manipulation associated with less protest [@Harvey2020]. In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006]. Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation. In sum, election manipulation can be performed by strong regimes for reasons that go beyond 'mere winning.'

<!--  One critique of this view, although still within the information/strategic framework, comes from Simpser's [-@Simpser2011] pathbreaking book. Simpser argues that when wanipulation efforts are excessive---well beyond what is needed simply to win---they demonstrate the organizational capacity, resources, and manpower of the ruling party [@Simpser2013]. Large-scale manipulation is, thus, a signal of strength rather than weakness, and can deter protest. By contrast, partial manipulation efforts that fail to secure a large majority vote-share signals the government lacks both popular support and mobilizational capacity, making opposition protest more likely. -->

A second school of thought has emphasized the ways in which biased elections may predispose individuals to engage in protest [@kuntz_more_2009]. In this view, election manipulation acts as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009]. Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991]. This conception is closely connected to the idea that manipulated elections carry legitimacy costs for the ruling party [@birch2011].

For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021]. This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].

The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance. A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b]. Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013]. Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006]. These studies suggest that electoral protest can occur at low or high levels of fraud, contingent on other factors\textemdash including subjective, and perhaps mistaken, beliefs.

The strategic and social-psychological explanations are, of course, interconnected. As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something. Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304]. Several empirical studies have, accordingly, found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014;@Harvey2020;@Rod2019].

There are thus unresolved tensions in the literature on the relationship between electoral manipulation and electoral protest. One branch of the literature argues that the risk of mass protest reduces incumbents' incentive to engage in manipulation, resulting in cleaner elections. A second branch argues that mass protest hinges on citizens' subjective assessments of election integrity, their individual grievances, and whether or not the election is seen to be stolen.  This view complicates the deterrence logic, since incumbents may minimize protest risk (and thus more freely manipulate) by reducing citizen grievances (i.e. intervening in the business cycle at election time, distributing club goods, etc.) or by controlling the information environment. Since citizens' subjective views of manipulation may diverge from the true level, incumbents may also face an incentive to manipulate the election severely in the hopes of achieving a large margin of victory, under the assumption that protest is likely even at low levels of true manipulation. Finally, while a few empirical studies do show evidence for the deterrence model, overall empirical support for the proposition is mixed at best and lacks a general test.

Finally, despite widespread election manipulation, electoral protest is relatively rare; even in cases where it does occur, most efforts to overturn fraudulent election results through protest have failed [@kalandadze2009electoral]. The apparent failure of protest risk to deter election fraud underlines our limited understanding of non-democratic ruling parties' decision-making. Below, it will be argued that election manipulation by its nature will rarely be associated with mass post-election protest. Elections in non-democracies are *generally* risky for incumbents, regardless of how free and fair they may be, as they create an opportunity for opposition collective action [@Knutsen2017] and reveal information about the government and society [@Pop-Eleches2015]. In most cases, however, election manipulation adds little to this risk.

# Theory

The foregoing literature review sketches out prior models of the relationship between election manipulation and protest. That existing work has several implications that will be tested here. First, does a higher risk of protest deter subsequent election fraud? And second, do high levels of fraud increase the risk of subsequent protest?

## Predictions from prior work: Fraud and the risk of protest initiation

As noted above, much of the formal literature on election manipulation assumes that the risk of protest serves as a deterrent to election manipulation. In addition, the logic of deterrence suggests that increasingly severe election fraud should make post-election more likely, either unconditionally or when the incumbent's margin of victory is low. More extensive election-day manipulation, in these models, is generally taken as a signal of regime unpopularity. This in turn makes citizens [@Egorov2021;@Little2015] or opposition parties [@Luo2020;@Rozenas2016b;@magaloni2010] more likely to stage protests. Previous research has also suggested that a fraudulent election is more likely to trigger protest when social grievances are high [@Brancati2016;@tucker2007]. With some exceptions, such as the study by Lankina and Skovoroda [-@Lankina2017], most research on election fraud and protest treats the latter as a binary outcome\textemdash protests either occur or not. This leads to the following hypotheses:

>H1: Greater pre-election protest risk will be associated with lower levels of election-day manipulation.

>H2: The likelihood of protest initiation will increase as the severity of election fraud increases.

>H3: The effect of election manipulation on the likelihood of protest initiation will increase as the incumbent's margin of victory falls.

>H4: The effect of election manipulation on the likelihood of protest initiation will increase as economic grievances rise.

However, there are reasons to doubt the veracity of the deterrence logic. Election fraud and other forms of manipulation that do not directly mobilize voters (such as vote-buying) cannot be a reliable indicator of the strength or weakness of the regime; consequently the 'signal' provided by a fraudulent election will not be able to reliably generate strategic behavior by the opposition. The central problem is that election manipulation is a tool that can be used by weak and strong governments alike [@simpser2013]. A strong government may allow for relatively clean elections, in order to gather more information about the opposition and society [@ananyevDictatorsSignalStrength2022;@trinhTeaLeafElections2022]; a weak government may be unable to muster the necessary resources [@greene_why_2007]. The observation of a relatively clean election alone does not discriminate between the two. Likewise, a strong government may manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver]. This is all in contrast to the typical deterrence model, in which a weaker government attempts to manipulate in order to obscure its weakness, but in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are weakly associated with more election fraud. For these reasons, acts of manipulation can be interpreted either as signs of regime strength or weakness [@Pop-Eleches2015]. It is also in line with the formal model proposed by Little [-@Little2012], in which citizens\textemdash knowing the regime is capable of manipulating the results\textemdash discount an expected level of fraud from the results when deciding whether or not to protest.

Additionally, in-depth investigations of the causes of the Color Revolutions (in particular) have emphasized the role of opposition agency and learning in staging post-election protest. In this view, the likelihood of mass electoral protest is increased by deliberate efforts by opposition parties to form a unified opposition coalition, to raise popular expectation of a stolen election, to prepare logistics, and to actively monitor the vote and publicize discrepancies [@bunce2010]. Others have noted structural factors like the importance of ties to democratic Western states and the size and strength of the ruling party and security apparatus in influencing the course of the Color Revolutions [@LucanWay2008]. In either case, these factors create an incentive for opposition protest whether electoral fraud is minor or severe\textemdash a kind of case of 'crying wolf'\textemdash an outcome that is also posited in some of the formal modeling literature [@Little2015]. As a result, it is likely that the severity of fraud has little bearing on protest risk; what matters is the underlying balance of strength between the incumbent regime and its opposition.

## The course of protest: do higher levels of fraud make protest waves more significant?

While it is useful to check prior predictions on protest initiation, by emphasizing binary protest outcomes at the expense of trends over time, prior work has overemphasized the risk of popular backlash to incumbents. While many incumbents may indeed prefer a quiet post-election period, the occurrence of protest can also have regime-sustaining benefits. When unexpected post-election protests occur, trust in the government can increase, especially among opposition supporters [@Frye2019]. As the number of protests increases, support for the protesters' aims may fall; this effect is even larger for those who consume pro-government media [@tertytchnayaElectoralProtestsPolitical2020]. These empirical results are in line with models proposed by Lohmann [-@lohmannDynamicsInformationalCascades1994] and Shadmehr and Bernhardt [-@shadmehrCollectiveActionUncertain2011]; they argue that when some citizens are too predisposed to protest, others become more inclined to stay home\textemdash dooming the first group's efforts to failure. Such influences help offset the 'risk' of protest to incumbents, especially when protest movements are relatively small and/or isolated.

What incumbents likely *do* seek to avoid are large, widespread, and long-lasting protest movements that strain the government's resources and put pressure on elites [@chenowethPhysicsDissentEffects2019]. Here I argue that, once protests begin, the extent of electoral fraud is unlikely to affect their scope. The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014]. That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].

In many cases, the government doubles down on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007]. Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022]. In other cases, the government may attempt to accommodate the opposition, by making concessions on policy or institutions [@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979]. Alternatively, concessions may be aimed at the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019]. In addition, governments may learn from the experiences of other governments, just as opposition groups may learn from opposition. Anticipating threats from below may encourage less democratic governments to invest more in the security services, or to reform political and legal institutions in order to stifle potential protests before they take shape [@hallAuthoritarianLearningConceptual2017a;@Morgenbesser2020].

What results is a test of strength between the anti-incumbent protest coalition and the regime with its supporters [@mcadamDynamicsContention2001;@tillyRegimesRepertoires2010]. In essence, as Beaulieu [@Beaulieu2014a] notes, the onset of electoral protest marks a case of bargaining failure. Just as in the international context, the ensuing conflict acts to reveal information on the strength and resolve of the two sides.  In such a contest, the incumbent enjoys multiple advantages, though it may lack the support of a majority of the population. It controls the resources of the state, and can call upon the support of those elites who benefit from the status quo. High-capacity states, as a result, are more likely to be able to engage in successful repression [@tarrowPowerMovementSocial2011].

Though a biased election may temporarily allow diverse opposition groups to set aide their differences [@trejo_ballot_2014], this internal diversity can enable the incumbent to divide the opposition with selective repression or concessions. While the regime may have its own splits---between reformists and hardliners, for example---the emergence of protest can act to solidify support for the ruling party among its supporters. This can be especially true non-democracies where elites have strong incentives to hold to power, in the form of access to rents and preservation of immunity for prosecution   [@baturoDemocracyDictatorshipTerm2014].

In this struggle between the potential protest coalition and the incumbent government, protesters and their possible allies behave as "passionate economists" whose balance of emotional versus strategic considerations can be updated over time [@VanZomeren2012]. What is the likelihood that protest action can help resolve the grievance, and does the benefit of participating outweigh the risk of personal consequences to the individual for participating?  A fraudulent election may heighten citizens' grievances and generate uncertainty around the underlying popularity of the regime, creating a moment of elevated protest risk. Crucially, though, this evaluation is dynamic and can be updated in response to new information and new events [@VanZomeren2012]. If protest movements have sufficient numbers and organizational resources to put the regime on the back foot, this can lead to more numerous protests as individuals' sense of group efficacy grows and their sense of risk declines [@Chenoweth_Stephan2011]. On the other hand, if the regime is well-resourced, capable, and unified, its actions to end a protest movement can lead to a negative re-evaluation of the movement's odds of success. This does not, of course, eliminate citizens' sense of grievance. Instead, it leads them to abandon 'approach-oriented' forms of coping, to avoidance-oriented forms: demobilization and disengagement [@van_zomeren_put_2004;@Young2019]. For example, while ruling-party supporters in Russia disapprove of election fraud [@Reuter2021], they are more likely to withdraw from politics after exposure to electoral protest than to support the opposition [@Tertytchnaya2020].

Here, this logic will be tested using two measures, both of which may indicate the strength of a protest movement, as well as the cost of protest to incumbent governments: the total number of protest events and the estimated number of protesters on the largest single day of protests. I assume that more numerous protests, and those that built to a higher peak participation, are costlier for governments to contain, all else held equal. If fraudulent elections reveal information about regime weakness or generate large grievance-motivated protest, more fraudulent elections should be associated with larger and more numerous protest waves. Prior theory expects these effects to be largest when margins of victory are smaller, or when economic grievances are high.

In summary, fraudulent elections are not likely to pose a significant protest risk to incumbents. While opposition actors may seize upon fraudulent elections to initiate a protest, the relationship between this decision and the level of fraud is likely to be weak. Opposition groups are also incentivized to stage protests regardless of the level of fraud, given the highly salient moment afforded by the election itself [@cunhaSignalReceivedAuthoritarian2022]. Fraud itself provides little information about regime strength, and even if a protest movement begins after the election, any useful information provided by the level of fraud is quickly replaced by signals from the protests themselves [@lohmannDynamicsInformationalCascades1994]. Protesters, bystanders, and elites can update their estimation of the relative strength of the protest coalition and the incumbent government by observing the size of crowds, the number of protests, and the government's reaction. Thus, I expect that Hypothesis 5\textemdash which might be assumed under other manipulation-oriented models of protest\textemdash to *not* be supported.

>H5: The number and maximum size of protest events are likely to increase as the severity of election fraud increases.

Once protests begin, the number and size of protest events is likely to be influenced by the organizational capacity of civil society and the opposition on the one hand, and on the ability of regime elites to remain united on the other. When opposition groups and civil society capacity are more organized and better resourced, it is likely that more protesters will take to the streets in more numerous events (see Chenoweth et al [-@chenowethStateRepressionNonviolent2017] for a review of factors making protest movements more resilient). By contrast, when the government appears well-resourced and resolved, more 'passionate economists' should stay home. Much of this balance is mediated by political and social elites' decision(s) to remain supportive of the incumbent; protests can create division among the incumbent elite (e.g. among hardliners and softliners [@schmitter1986transitions]) and incentives for disaffected elites to move against the incumbent leadership [@kimPathways2019]. In turn, the opportunity created by a divided elite\textemdash and their resources [@robertson_protest_2010]\textemdash can keep protest movements alive.

Since information on elite and mass preferences can be difficult to observe outside liberal democracies, in part due to preference falsification [@Kuran1991], it is difficult to measure the risk of elite splits *a priori*. However, Hale [-@hale_patronal_2014] makes the case that elites become incentivized to turn against incumbent governments when they expect the chief executive's time in office has become limited. While the incumbent's age and popularity may be signals of such, a constitutional term limit can also serve this role. As such, Hypothesis 6 predicts that protest number and size will increase for elections where the incumbent faces a term limit. This hypothesis serves as useful counter-point to the traditional view on election manipulation. Unlike election fraud (as presented here), a constitutional term limit creates a clear, unambiguous signal that enables coordination among protesters and elites that can influence the course of protest.

>H6: The number of protest events and the maximum event size are likely to increase following elections where the chief executive is.

To further test the mechanism outlined in the theory presented here\textemdash that election manipulation offers few clues to regime strength, and thus the course of electoral protests are determined by more structural factors\textemdash I examine the relationship between fraud and the survival of protest movements in cases where protests began but the incumbent eventually survived. Restricting the sample to these cases allows for a consistent interpretation of results: longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime, while shorter-lived protest movements are those which were more easily defeated.  In this sub-sample, I would first expect the relationship between election fraud and protest survival to be non-linear with respect to time\textemdash the longer the protest movement survives, the less relevant the fraud signal (such as it is) becomes). Second, it is expected that more severe election fraud will *not* be associated with longer protest waves, as would be predicted if fraud were an indicator of state weakness. Instead, structural factors should again be more influential; indicators of opposition and civil-society strength should help keep protest waves alive.

<!-- -Democracy leads to less repression: Bueno de Mesquita et al., 2005). Davenport & Armstrong (2004),  -->
<!-- -Prior repression makes repression more likely (Davenport 2007 (Tyrannical peace)) -->
<!-- -Resource wealth makes authoritarian survival more likely (Girod et al 2016) -->

<!-- From E-F 2013: Nevertheless, violent repression may have pernicious consequences, since it may bring about a -->
<!-- violent escalation or a radicalization of positions, thereby unleashing a violent response from some -->
<!-- groups and even potentially provoking a coup (Gurr, 1970; Ortiz, 2007; Pierskalla, 2010). Thus, -->
<!-- Opp and Ruehl (1990) suggest that repression may lead to micro-mobilization processes leading to -->
<!-- further protest activities and unrest. Likewise, Ginkel and Smith suggest that ‘dissident activity is -->
<!-- more likely to be effective in motivating rebellious action under highly repressive conditions’ -->
<!-- (1999: 292). Some argue that when nonviolent actions are violently repressed, protesters may turn -->
<!-- to violent activities (Lichbach, 1987; Moore, 1998), as the Syrian case illustrates. Similarly, -->
<!-- Goodwin (2001) argues that revolutions are more likely in patrimonial and very repressive states. -->
<!-- For instance, in Iran on Black Friday in September 1978, the regime’s security forces killed hun- -->
<!-- dreds of anti-shah protesters in Tehran. After a brief decrease in activity, public dissent and strikes -->
<!-- quickly resumed and escalated. The mobilizations that took place in December gathered hundreds -->
<!-- of thousands of people calling for the shah’s rule to end. In January 1979, the shah left the -->
<!-- country -->

<!-- ...restrictions on civil rights can thwart the ability of some -->
<!-- actors (especially the opposition) to organize, coordinate, and mobilize in the short run -->

<!-- On uncertainty of repression and protest, from Girod et al: Yet the differences in demobilization of protest movements following repression remain largely unexplained. Government repression of protesters is strongly associated with quelling protest activity in some cases (Lyall, 2009; Schatzman, 2005), and triggering increased mass mobilization in others (Francisco, 1995; Khawaja, 1993; Mason and Krane, 1989). Complicating matters further, in some cases the relationship between repression and protest appears to change over time (Francisco, 1995; Inclan, 2009; Mason and Krane, 1989; Rasler, 1996; Tsebelis and Sprague, 1989), across levels of repression (DeNardo, 1985; Levitsky and Way, 2010; Muller and Weede, 1990), and based on the complex ways in which “threat” and “opportunity” combine to influence protest dynamics (Goldstone and Tilly, 2001). Because fine-grained global data on repression of mass movements have not been available until now (Davenport, 2005: viii; 2007a: 10), extant studies depend largely on only a few cases. In short, the literature has yet to identify cross-national patterns to explain when publics in an authoritarian regime respond to repression with greater collective action vs. disbanding. -->

 <!-- Chernykh and Svolik, in particular, argue that by revealing information about manipulation, even moderately independent courts can increase the risk of protest against manipulated elections (thus deterring manipulation), while also reducing the risk that opposition groups mistakenly protest clean elections. This result is in keeping with the argument that electoral protests are the outcome of bargaining failure among incumbents and opposition groups, which are worsened in low-information environments [@Beaulieu2014a]. -->

<!-- However, one of the central functions of independent judiciaries outside of liberal democracies is to create mechanisms through which citizens grievances can be addressed [@Whiting2017; @Solomon2007; @Hilbink2007d]. Relatively impartial courts create an opportunity for citizens to resolve grievances against the state and against other individuals who may be more powerful or influential. For example, survey-experimental data show that citizens of several Middle Eastern non-democracies were more willing to engage in legal action to address their grievances when they believe courts are more independent and assertive [@Schaaf2021]. Moreover, the belief that procedures are generally followed impartially can bolster the state's claim to legitimacy [@Gerschewski2013], increasing citizens' trust in authority and willingness to abide by political outcomes [@Grimes2006]. Evidence from European democracies suggests that that de facto independence is key for such judgments [@Buhlmann2011]. Taken together, greater de facto judicial independence ought to reduce protest risk, both by creating a channel for the state to address citizens' general grievances and by offering an alternative venue to the street for addressing specific electoral concerns.  -->

<!-- These two functions for courts are at odds with regard to electoral disputes. As a result, the relationship between judicial independence and post-election protest is conceptually and empirically muddled---despite the centrality of protest risk to predominant models of electoral manipulation. The formal model proposed by Chernykh and Svolik [-@Chernykh2015] is not tested empirically, while Chernykh [-@chernykh2014] and Eisenstadt [-@eisenstadt_courting_2003] both find no effect for judicial independence on the likelihood of protest in post-communist Europe and Mexico, respectively. This calls into question the prediction that more independent courts make protest more likely in the presence of election manipulation, though to my knowledge the question has not been studied in a large-n cross-national context. -->

@@

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-30.csv"))

@@

#Margins with updates

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

@@

nelda2.sub <- nelda2.sub %>% filter(v2x_regime <= 2) #To-do: Extent to include liberal dems in appendix

@@

These hypotheses are tested using data from three datasets: NELDA [@hyde_2012_marinov], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019]. The unit of observation for this study is an election period as recorded in NELDA. This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.  I limit the sample by excluding closed autocracies (where no national elections are held). Since the theory of popular backlash to manipulation has largely been articulated in an electoral-democratic context, I exclude liberal democracies from the main analysis.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1 or 2.] As a robustness check for this sampling decision, I also run the main models with data including liberal democracies in the appendix, with no substantial changes in the outcomes. The first set of models, focusing on Hypothesis 1, include all the remaining cases in the dataset. For the subsequent models, focusing on post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud. Finally, the beginning of the date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012. The resulting dataset includes 1110 total election-period observations across 118 countries.

## Dependent variables and modeling strategies

>>>>

On August 9, 2020, Belarus held an election in which President Aleksander Lukashenko ran for a sixth term. The official results credited Lukashenko with over 80\% of the vote, a result which observers regarded as wildly inflated [@GolosPlatformPresents2020]. In response, hundreds of thousands of Belarusians took to the streets over the course of several months. Though the protests put significant strain on the regime, a harsh campaign of repression (and support from neighboring Russia) ultimately quelled the protests; Lukashenko was inaugurated into a new five-year term.

The Belarusian case is typical in some ways, and an outlier in others. Like many previous elections in Belarus, and in authoritarian regimes around the world, the contest was heavily managed and the results inflated by fraud. However, the sustained, massive protest wave that occurred post-election is extremely unusual. However, the regime's ultimate victory over the protest movement\textemdash while costly and never guaranteed\textemdash is also the typical outcome of post-election protest.

These patterns highlight a puzzle generated by existing research on the relationship between election manipulation and popular protest. The topic has significant implications for authoritarian governance and democratization, and as such has been a very active subject of study. However, the various theoretical perspectives brought to bear on the topic have generated competing predictions. To briefly summarize: for some models, election fraud is a tool of weak incumbents. As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon2011]. For others, fraud is committed by weak and strong incumbents alike, making protest a risky gamble the opposition undertakes for other reasons. In these models, fraud may or may not be deterred based on other exogenous factors: a dubious, though not completely dismissive, prospect for democratization and self-enforcing democracy. Lastly, a third school of thought holds that election fraud is an indicator of genuine incumbent strength. This approach, which is expanded upon below,

@@

Testing this proposition requires looking beyond binary measures of protest, and toward measures of its scope. Research on the risk posed to governments by electoral protest often acknowledges---whether explicitly or implicitly---the importance of broad-based protest campaigns, rather than one-off events [@Bunce2006;@tucker2007;@DAnieri2006;@Carnaghan2016;@bunce2010]. This implicit assumption has largely been overlooked in theories and empirical studies of the relationship between election manipulation and protest, however. In most cases, the risk of protest has both been conceptualized and operationalized as a binary outcome: protests occur or they do not. This approach obscures the importance of the number and size of electoral protests in a ruling party's calculation of the risk involved in manipulation an election.

Using cross-national data on 1110 elections from 1989 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019], this paper finds little support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs. Furthermore, it shows that more severe election fraud is not associated with more numerous or larger post-election protests. Instead, these protest dynamics appear driven more by structural and institutional factors, including in particular whether or not the incumbent chief executive is term limited.

Since fraud is not associated with the overall number or size of of events in a protest wave, the risk to governments has likely been overemphasized. That is, even if protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them. Once protests begin, the noisy signal of regime strength provided by the election is replaced by a very clear signal: the force of the regime's response to the emerging protest movement in contest with the mobilizational capacity of the opposition.  To the extent that larger protest waves are more damaging to the government than small ones, protest risk does not appear to be a substantial constraint on the severity of election fraud. Troublingly, these findings indicate that ruling parties should follow an 'all-in' strategy on election manipulation. More manipulation is not associated with a larger number of protests, but a large margin of victory is associated with a lower risk of protest initiation. These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory. They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.

@@

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-01-16.csv"))

@@

@@

nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1) #Includes only authoritarian

@@
